The importance of a complete description of dependence between variables for ecology and related fields

### 2019-05-09 Paper.Rmd
small changes
the commit right befor this one was the start of making changes to respond to Lawrence's comments and those of John Drake after the preproposal

updated on `r Sys.Date()`

@@

\noindent A complete statistical description of dependence between variables for ecology and related fields

@@

<!--summary table with info of all data used in Paper-->
```{r tab_summary_data_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", "Cedar Creek data", 
                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances","Methane flux")
tab_data_info$Category<-c("Environmental","Species-trait","Species-trait","Community level",
                          "Population","Phenological","Population","Ecosystem functioning")
tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
                                               paste(nrow(d_birdsBMR)," birds",sep=""),
                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
                                               "13 locations with atleast 50 dates of data")
tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")

knitr::kable(tab_data_info, 
             format="latex", align="l",
             caption = "Summary table for the data we used. Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="2.8 cm")%>%
             row_spec(5:8, bold=T)
```

>>>>

@@

The importance of a complete copula description of dependence between variables for ecology and related fields

Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed with copulas

Complex, informative, and previously ignored dependencies propogate throughout ecology and are revealed using copulas

Copulas reveal complex, informative, and previously ignored dependence
structures between variables across ecological disciplines

Copulas reveal complex, informative, and previously ignored dependence
structures which in turn illuminate multiple ecological disciplines

Copulas reveal complex, informative, and previously ignored dependence
structures which in turn can illuminate every subdiscipline of ecology

Copulas reveal complex, informative, and previously ignored dependence
structures which in turn can illuminate every subfield of ecology

Copulas reveal complex, informative and previously ignored dependence
structure between variables throughout ecology

The complexity of dependencies between ecological variables, previously
ignored, is revealed by copulas and in turn illuminates multiple 
disciplines of ecology

@@

<!--summary table with info of all data used in Paper-->
```{r tab_summary_data_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane flux")
tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
                          "Population","Phenological","Population","Community level","Ecosystem functioning")
tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
                                               paste(nrow(d_birdsBMR)," birds",sep=""),
                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                               "13 locations with atleast 50 dates of data")
tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")

knitr::kable(tab_data_info, 
             format="latex", align="l",
             caption = "Summary table for the data we used. Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="2.8 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```
